On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit

### 2022-07-22 Manuscript/manuscript.tex

\newcommand{\jy}[1]{\textcolor{blue}{JY: #1}}

@@

The Kolmogorov-Smirnov (KS) test is one of the most popular goodness-of-fit 
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample from some continuous distribution. 
The null hypothesis $H_0$ is that the population distribution is $F$.
Let $F_n(\cdot)$ be the empirical cumulative distribution. The KS statistic is
\begin{equation}
  \label{eq:ks_standard}
  D = \sup_x | F_{n}(x) - F(x) |.
\end{equation}
The distribution of $D$ under $H_0$ turns out to be independent of the
distribution $F$ and a table of critical values for $D$ has been constructed
\citep{Massey} for various sample sizes $n$ and significance 
levels $\alpha$. If the value of $D$ exceeds the test's corresponding critical 
value the null hypothesis is rejected. The KS test is available in popular 
statistical software packages, such as function \texttt{ks.test} in R 
\citep{R, Marsaglia}.

@@

\citet{Noether} provides a brief demonstration of the 
conservative character of the KS test when using discrete functions. 
There are remedies that allow the KS test to be used in discrete cases. 
\citet{Conover} describes a method for finding the exact critical level for 
discrete cases and derives its power. \citet{Gleser} provides methods for 
adapting existing algorithms to provide results when $F(x)$ is discontinuous.
The discrete KS test is available in various statistical software packages.
The package \textsf{dgof} implements the methods from \citet{Conover} for 
one-sided tests and the methods from \citet{Gleser} for exact two-sided 
p-values \citep{Arnold}. The \textsf{KSgeneral} package provides a fast and 
accurate method for when $F(x)$ is arbitrary, discontinuous, or continuous that 
uses an alternative fast Fourier transform based method \citep{Dimitrova}.
This issue with discrete or rounded is not the focus of this paper, though is a
common misuse that should be noted.

The standard KS test is not applicable when the hypothesized distribution 
contains fitted parameters. \citet{Steinskog} ``discovers'' the change in power 
when using estimated parameters and stresses caution in using the KS test in 
such ways. \citet{Lilliefors} shows for the normal distribution that using the 
standard table when values of the mean and standard deviation are estimated 
obtains extremely conservative results. This is supported by \citet{Capasso}, 
who concludes that failing to re-estimate the parameters may lead to wrong, 
overly-conservative approximations to the distributions of goodness-of-fit test 
statistics based on the empirical distribution function. \citet{Capasso} also 
notes that the impact of this mistake may turn out to be dramatic and does not 
vanish as the sample size increases. Remedies are provided by \citet{Babu} and 
\citet{Genest} in the form of bootstrap. \citet{Babu} details the bootstrap
procedure for goodness-of-fit tests and notes that both parametric and 
non-parametric procedures lead to correct asymptotic levels, however there is a 
correction required for the non-parametric case. \citet{Genest} provides 
validity for using parametric bootstrap with various goodness-of-fit tests.

In the case of serially dependent data, \citet{Durilleul} demonstrates that the 
KS statistic is too liberal for medium-to-high positive autocorrelation values. 
\citet{Durilleul} also shows that for negative autocorrelation values, the 
behavior is asymmetrical with respect to positive values. For remedies, 
\citet{Weiss} provides a procedure that is applicable specifically for data 
modeled by the second-order auto-regressive (AR) process where the parameters 
are known. \citet{Lanzante} tests various strategies for dealing with temporal 
dependence and concludes that a test based on Monte-Carlo simulations performed 
the best. We propose a parametric bootstrap procedure involving copulas to 
account for dependence.

@@

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of 
the empirical distribution is serially dependent. A bootstrap procedure 
employing copulas is a working solution. Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs. The 
copula procedure can be adjusted for fitted parameters to be available as a fix. 
Section~\ref{sec:conclusion} concludes with a discussion.

@@

The null distribution of the KS statistic changes when the hypothesized
distribution contains fitted parameters. In this scenario, the null hypothesis
is $H_0$: the random sample $X_1, \ldots, X_n$ comes from a continuous
distribution $F_{\theta}$ with unspecified parameter vector $\theta$.
Let $\hat\theta_n$ be an estimator of $\theta$, which could be, for example,
maximum likelihood estimator of moment estimator. The test statistic is
\begin{equation}
  \label{eq:ks_fitted}
  D = \sup_x | F_n(x) - F_{\hat\theta_n}(x) |.
\end{equation}
Since $F_{\hat\theta_n}$ is not the same as the true data generating
$F_\theta$, the null distribution of $D$ obtained in existing implementations in
software packages which assumes completely known $F_\theta$ no longer applies
\citep{Steinskog}.

To demonstrate the consequences of this problem, a simulation is performed. A 
random sample $X_1, \ldots, X_n$ is generated from the standard normal 
distribution with sample size $n = 100$. The p-values of $1000$ replicate tests 
are displayed in the Naive plot of Figure~\ref{fig:hist_fitted}. Each KS test 
was performed using fitted parameters, i.e. the hypothesized distribution is 
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample 
variance. Since the data is generated from a standard normal distribution with 
seemingly all assumptions met, a uniform distribution of $U(0, 1)$ is expected 
for the p-values. However this is under the assumption that the KS test
assumptions hold, which no longer do due to using fitted parameters. Therefore, 
there is notable deviation from the uniform distribution.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_fitted}
  \caption{The Naive plot (left) is the histogram of p-values from the standard 
  KS test with fitted parameters. The Parametric plot (middle) is the histogram
  of p-values from implementing parametric bootstrap. The Babu plot (right) is 
  the histogram of p-values from implementing nonparametric bootstrap with a 
  correction for bias. In each case, $1000$ replicate tests were performed on 
  the standard normal distribution with sample size $n = 100$. $B = 1000$ 
  bootstrap samples are obtained for each test.}
  \label{fig:hist_fitted}
\end{figure}

@@

The KS test also displays issues in the case of dependent data. As mentioned, an 
assumption of the test is that the data is independent. Unfortunately, real data
is often temporally or spatially dependent and the results of a goodness-of-fit 
test would be valuable. When the KS test is performed on dependent data it 
performs poorly. In this situation, the null hypothesis is $H_0$: the random 
sample $X_1, \ldots, X_n$ come from a continuous distribution $F$ where $F$ has 
completely specified parameters. The test statistic is the same as 
Equation~\eqref{eq:ks_standard}. However, since the data is dependent, the null 
hypothesis is wrong and must be corrected. This is demonstrated with a 
simulation in Figure~\ref{fig:hist_correlation}. Data is generated from a 
first-order autoregressive model (AR(1)) with a standard normal distribution. 
P-values are gathered from $1000$ replicate tests for different levels of the 
AR coefficient $\psi$, varying from $(-3,3)$. In the presence of serially 
dependent data, the distribution of p-values no longer follows $U(0, 1)$ as 
would be expected of a valid test. The results echo those of \citet{Durilleul} 
that the KS statistic is too liberal for positive autocorrelation values, and 
that the behavior is asymmetrical for negative values.

@@

In order to correct this, we can employ a parametric bootstrap procedure which
assumes a working serial dependence structure through copulas. A copula is a
multivariate distribution with standard uniform marginal distributions, which
completely characterizes the dependence structure of a multivariate
distribution \citep{Copula, Hofert}. For simplicity, we assume a normal copula 
with an AR(1) structure to characterize the serial dependence of the 
observations. The AR(1) parameter $r$ of the normal copula is set to match the 
be the sample serial Spearman's rho of the observed sample. The procedure is as 
follows.

@@

To further demonstrate the effectiveness of the procedure, we perform similar 
simulations on other families of distributions. 
Figure~\ref{fig:hist_gamma_gev_D} displays the results of tests using dependent 
data generated from the gamma and the generalized extreme value distributions.
The procedure corrects the distribution of p-values for both distributions and 
is shown to be applicable irregardless of the family of distribution tested.

@@

Similar to Section~\ref{sec:dependence}, the copula fix is not a complete 
solution. Regardless of the true dependence in the data we assume an AR(1) 
dependence structure by taking the lag-1 sample auto-spearman rho. However, we 
can show that as long as the AR(1) assumption is a close approximation, the 
correction still provides a reasonable approximation that can be useful for 
practical purposes. Figure~\ref{fig:hist_ma1_arma_ar2_FD} shows the results of the 
procedure performed on data generated with dependence structures of  MA(1), 
ARMA(1, 1), and AR(2). Naively fitting parameters and not adjusting for 
dependence clearly deviates from a uniform distribution of p-values. Parametric 
bootstrap provides some correction but does not account for dependence, so 
therefore the results of the copula remedy are more accurate and favorable.
In the case of MA(1) and ARMA(1, 1), the true dependence appears close enough 
to our assumption of AR(1) that the results are reasonable. AR(2) however
appears to be just far enough from our assumption, showing a limitation of the 
procedure.

@@

It is also possible to apply our procedure to other families of distributions. 
The data in Figure~\ref{fig:hist_gamma_gev_FD} is generated from the gamma and 
generalized extreme value distributions. The results are favorable and show that 
the procedure outlined as a fix for fitted parameters and spatially dependent 
can be adapted for various families of distributions.

@@

The KS test has base assumptions that the hypothesized distribution is 
completely specified and the data is independent. When these assumptions are 
violated, the test is no longer accurate and remedies must be performed. In the 
case of fitted parameters, parametric and non-parametric bootstrap can restore 
the size of the test. A bias correction is required if the non-parametric form 
is used \citep{Babu}. In the case of dependent data, a procedure using bootstrap 
with copulas to model dependency shows positive results. When both assumptions 
are violated, that is the case where the data has dependence and parameters must 
be fitted, an adjusted copula procedure also shows positive results. The tests 
appear effective for a variety of families of distributions. The copula remedy 
is not a complete solution and has limitations. Regardless of the true 
dependence, we assume an AR(1) dependence structure. Therefore, if the AR(1) 
dependence structure is a close approximation of the truth, the fix can work. 
However, in cases such as AR(2), if the approximation is too far from the true 
dependence structure the fix does not completely solve the issue. As well, tests 
were only performed with the normal copula. It is possible that other copulas
could provide stronger results based on the true dependence of the data.

>>>>

\newcommand{\jy}[1]{\textcolor{blue}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{red}{EDS: (#1)}}

@@

The Kolmogorov-Smirnov (KS) test is one of the most popular goodness-of-fit 
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample from some continuous distribution. 
The null hypothesis $H_0$ is that the population distribution is $F$.
Let $F_n(\cdot)$ be the empirical cumulative distribution \eds{define ecdf}. 
The KS statistic \eds{cite} is
\begin{equation}
  \label{eq:ks_standard}
  D = \sup_x | F_{n}(x) - F(x) |.
\end{equation}
The distribution of $D$ under $H_0$ turns out to be independent of the
distribution $F$ and a table of critical values for $D$ has been constructed
\citep{Massey} for various sample sizes $n$ and significance 
levels $\alpha$. If the value of $D$ exceeds the test's corresponding critical 
value the null hypothesis is rejected. The KS test is available in popular 
statistical software packages, such as function \texttt{ks.test} in R 
\citep{R, Marsaglia}.

@@

\citet{Noether} provides a brief demonstration of the 
conservative character of the KS test when using discrete functions. 
There are remedies that allow the KS test to be used in discrete cases. 
\citet{Conover} describes a method for finding the exact critical level for 
discrete cases and derives its power. \citet{Gleser} provides methods for 
adapting existing algorithms to provide results when $F(x)$ is discontinuous.
The discrete KS test is available in various statistical software packages.
The package \textsf{dgof} implements the methods from \citet{Conover} for 
one-sided tests and the methods from \citet{Gleser} for exact two-sided 
p-values \citep{Arnold}. The \textsf{KSgeneral} package provides a fast and 
accurate method for when $F(x)$ is arbitrary, discontinuous, or continuous that 
uses an alternative fast Fourier transform based method \citep{Dimitrova}.
This issue with discrete or rounded data is not the focus of this paper, though
 it is a
common misuse that should be noted.

The standard KS test is not applicable when the hypothesized distribution 
contains fitted parameters. \citet{Steinskog} ``discovers'' the change in power 
when using estimated parameters and stresses caution in using the KS test in 
such ways. \citet{Lilliefors} shows for the normal distribution that using the 
standard table when values of the mean and standard deviation are estimated 
yields extremely conservative results. This is supported by \citet{Capasso}, 
who concludes that failing to re-estimate the parameters may lead to wrong, 
overly-conservative approximations to the distributions of goodness-of-fit test 
statistics based on the empirical distribution function. \citet{Capasso} also 
notes that the impact of this mistake may turn out to be dramatic and does not 
vanish as the sample size increases. Remedies are provided by \citet{Babu} and 
\citet{Genest} in the form of bootstrap. \citet{Babu} details the bootstrap
procedure for goodness-of-fit tests and notes that both parametric and 
non-parametric procedures lead to correct asymptotic levels, however there is a 
correction required for the non-parametric case. \citet{Genest} provides 
validity for using parametric bootstrap with various goodness-of-fit tests.

In the case of serially dependent data, \citet{Durilleul} demonstrates that the 
KS statistic is too liberal for medium-to-high positive autocorrelation values. 
\citet{Durilleul} also shows that for negative autocorrelation values, the 
behavior is asymmetrical with respect to positive values. For remedies, 
\citet{Weiss} provides a procedure that is applicable specifically for data 
modeled by the second-order auto-regressive (AR) process where the parameters 
are known. \citet{Lanzante} tests various strategies for dealing with temporal 
dependence and concludes that a test based on Monte-Carlo simulations performed 
the best. %We propose a parametric bootstrap procedure involving copulas to 
%account for dependence.

@@

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of 
the empirical distribution is serially dependent. A bootstrap procedure 
employing copulas to account for dependence is proposed as a working solution. 
Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs. The 
copula procedure can be adjusted for the use of fitted parameters as a remedy. 
Section~\ref{sec:conclusion} concludes with a discussion.

@@

The null distribution of the KS statistic changes when the hypothesized
distribution contains fitted parameters. 
\eds{what is the null distribution when not using fitted parameters}.  
In this scenario, the null hypothesis
is $H_0$: the random sample $X_1, \ldots, X_n$ comes from a continuous
distribution $F_{\theta}$ with unspecified parameter vector $\theta$.
Let $\hat\theta_n$ be an estimator of $\theta$, which could be, for example, a
maximum likelihood estimator or moment estimator. The test statistic is
\begin{equation}
  \label{eq:ks_fitted}
  D = \sup_x | F_n(x) - F_{\hat\theta_n}(x) |.
\end{equation}
Since $F_{\hat\theta_n}$ is not the same as the true data generating
$F_\theta$, the null distribution of $D$ obtained in existing implementations in
software packages which assumes completely known $F_\theta$ no longer applies
\citep{Steinskog}.

To demonstrate the consequences of this problem, a simulation is performed. A 
random sample $X_1, \ldots, X_n$ is generated from the standard normal 
distribution with sample size $n = 100$. The p-values of $1000$ replicate tests
 of normality  
are displayed in the Naive plot of Figure~\ref{fig:hist_fitted}. 
\eds{how are these p-values calculated?}
Each KS test 
was performed using fitted parameters, i.e., the hypothesized distribution is 
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample 
variance. Since the data is truly generated from a standard normal distribution,
 %with 
%seemingly all assumptions met, 
a uniform distribution of $U(0, 1)$ would naively be expected for the p-values.
However this should only be expected when the KS test assumptions hold, and they
 do not due to using fitted parameters. Therefore, 
there is notable deviation from the uniform distribution.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_fitted}
  \caption{The Naive plot (left) is the histogram of p-values from the standard 
  KS test with fitted parameters. The Parametric plot (middle) is the histogram
  of p-values from implementing parametric bootstrap. The Babu plot (right) is 
  the histogram of p-values from implementing nonparametric bootstrap with a 
  correction for bias. In each case, $1000$ replicate tests were performed on 
  the standard normal distribution with sample size $n = 100$. $B = 1000$ 
  bootstrap samples are obtained for each test. \eds{perhaps use density hist vs
	freq. hist (here and elsewhere)?}}
  \label{fig:hist_fitted}
\end{figure}

@@

The KS test also displays issues in the case of dependent data. As mentioned, an 
assumption of the test is that the data is independent. Real data, however,
is often temporally or spatially dependent and the results of a goodness-of-fit 
test would be valuable. When the KS test is performed on dependent data it 
performs poorly. In this situation, the null hypothesis is $H_0$: the random 
sample $X_1, \ldots, X_n$ come from a continuous distribution $F$ where $F$ has 
completely specified parameters. The test statistic is the same as 
Equation~\eqref{eq:ks_standard}. However, since the data is dependent, the null 
hypothesis is wrong and must be corrected. 
\eds{null hypothesis or null distribution? or both?}
This is demonstrated with a 
simulation in Figure~\ref{fig:hist_correlation}. Data is generated from a 
first-order autoregressive model (AR(1)) with a standard normal distribution. 
P-values are gathered from $1000$ replicate tests for different levels of the 
AR coefficient $\psi$, varying from $(-3,3)$. 
\eds{how are these p-values calculated?}
In the presence of serially 
dependent data, the distribution of p-values no longer follows $U(0, 1)$ as 
would be expected of a valid test. The results echo those of \citet{Durilleul} 
that the KS statistic is too liberal in the presence of positive
autocorrelation, and that the behavior is asymmetrical for negative
 autocorrelation.

@@

In order to correct this, we can employ a parametric bootstrap procedure which
assumes a working serial dependence structure through copulas. A copula is a
multivariate distribution with standard uniform marginal distributions, which
completely characterizes the dependence structure of a multivariate
distribution \citep{Copula, Hofert}. For simplicity, we assume a normal copula 
with an AR(1) structure to characterize the serial dependence of the 
observations. The AR(1) parameter $r$ of the normal copula is set to match the 
sample serial Spearman's rho of the observed data. 
\eds{What is the difference between $\psi$ and $r$?}
The procedure is as 
follows.

@@

To further demonstrate the effectiveness of the procedure, we perform similar 
simulations on other families of distributions. 
Figure~\ref{fig:hist_gamma_gev_D} displays the results of tests using dependent 
data generated from the gamma and the generalized extreme value distributions.
The procedure corrects the distribution of p-values for both distributions and 
is shown to be applicable regardless of the family of distribution tested.

@@

Similar to Section~\ref{sec:dependence}, the copula approach is not a complete 
solution. Regardless of the true dependence in the data we assume an AR(1) 
dependence structure by taking the lag-1 sample auto-spearman rho. However, we 
can show that as long as the AR(1) assumption is a close approximation, the 
correction still provides a reasonable approximation that can be useful for 
practical purposes. Figure~\ref{fig:hist_ma1_arma_ar2_FD} shows the results of the 
procedure performed on data generated with dependence structures of  MA(1), 
ARMA(1, 1), and AR(2). Naively fitting parameters and not adjusting for 
dependence clearly deviates from a uniform distribution of p-values. Parametric 
bootstrap provides some correction but does not account for dependence, so 
therefore the results of the copula remedy are more accurate and favorable.
In the case of MA(1) and ARMA(1, 1), the true dependence appears close enough 
to our assumption of AR(1) that the results are reasonable. AR(2) however
appears to be just far enough from our assumption, showing a limitation of the 
procedure.

@@

It is also possible to apply our procedure to other families of distributions. 
The data in Figure~\ref{fig:hist_gamma_gev_FD} is generated from the gamma and 
generalized extreme value distributions. The results are favorable and show that 
the procedure outlined as a remedy for fitted parameters and spatially dependent 
can be adapted for various families of distributions.

@@

The KS test has base assumptions that the hypothesized distribution is 
completely specified and the data is independent. When these assumptions are 
violated, the test is no longer accurate and remedies must be performed. In the 
case of fitted parameters, parametric and non-parametric bootstrap can restore 
the size of the test. A bias correction is required if the non-parametric form 
is used \citep{Babu}. In the case of \eds{serially} dependent data, a procedure 
using bootstrap 
with copulas to model dependency shows positive results. When both assumptions 
are violated, i.e., where the data has \eds{serial} dependence and parameters 
must 
be fitted, an adjusted copula procedure also shows positive results. The tests 
appear effective for a variety of families of distributions. The copula remedy 
is not a complete solution and has limitations. Regardless of the true 
dependence, we assume an AR(1) dependence structure. Therefore, if the AR(1) 
dependence structure is a close approximation of the truth, the approach can work. 
However, in cases such as AR(2), if the approximation is too far from the true 
dependence structure the approach does not completely remedy the issue. As well, 
tests were only performed with the normal copula. It is possible that other 
copulas could provide stronger results based on the true dependence of the data.
